---
title: "Speaker Series"
author: "Juan Heslop & Quan Tran Minh"
date: "12/14/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(tidyverse)
library(forecast)
library(tsibble)
library(hts)
library(lubridate)
library(gridExtra)
library(usmap)
theme_set(theme_bw())

source("./functions.R")
```

# Superstore Sales & Profit 

Get and Preprocess the Data: 

```{r dates}
data = read.csv("./Data/superstore.csv")
data = preprocess(data)
prices = get_products(data)
# population = read.csv('./Data/population.csv')
```

Understanding the variables: 

- *Sales:* Total paid by customer (considers discounts as well). Same as Revenue
- *Profit:* Sales - Cost of Good - Operating and Other Expenses - Interest - Taxes

## Company Overview: 

  For confidienciality reasons, we could not have acces to the exact name of the company which provided witht his data. Therefore, time plots and variations cannos be specifically tied to real events: 

```{r}
d = aggr(data, 'Quarter')

sales = data.frame(Sales = c(sum(d[1:3, 1]),
sum(d[4:6, 1]),
sum(d[7:9, 1]),
sum(d[10:12, 1])))

profits = data.frame(Profits = c(sum(d[1:3, 2]),
sum(d[4:6, 2]),
sum(d[7:9, 2]),
sum(d[10:12, 2])))

net_margin = profits/sales
```

As has been said, our company is a small retail company. Being a small company, we are more dependent on external economic factors. We are less stable, less immersed in the market, and will surely suffer the consequences should there be an economic disaster. This makes our analysis even more relevant. 

```{r}
autoplot(ts(sales, start = 2016))
```

- Revenue: Steadily increasing from 300k per year to 500k in the previous 2019 year
- Profit: Also increased corresponging to an increase in sales, reaching a maximum of 70k for the final year

(These are very low numbers when compared with the retail giants in the USA, as Wallmart, that had a revenue of over 300 billion for 2019 in the USA alone)

These produce the following net margins for the fiscal years 2016 through 2019: 

```{r}
autoplot(ts(net_margin, start = 2016))
```

- How to optimize post covid. Everything is in hold. Starting again, how could we optimize retail business. 2020 is already covid, so it is out, it is an outlier. We postulate this scenario. 

The Average Net Margin is 12.47%. To compare this number within the industry, this must be filtered by the type of industry as it is very dependent. Our retail contains 3 subfields: 

```{r}
data_quarter = aggr(data, 'Quarter', 'Category')
round(apply(data_quarter, 2, sum)[4:6]/apply(data_quarter, 2, sum)[1:3], 3)
```

Furniture: 2.5%. Furniture stores sell high-end merchandise, which gives the impression that the margins are high. However, it’s costly to make quality furniture. This is why the net profit margin is so low. However, still it is pretty low comparing with competing companies in the industry given that the net profit margin for the industry is about 4% in average. 

Technology: 17.4%. Average of the industry is 17% so okay. https://www.investopedia.com/ask/answers/060215/what-average-profit-margin-company-telecommunications-sector.asp

Office Supplies: Well over the average value, doing very good

So we see a problem: the furniture market that we are investing on is not generating a significant margin that is increasing our overall sales, and not doing so competitively compared to other industries. This has problems, because given that ecnonomic circumstances might get worse from now on, investing in furniture might be risky. 

## See which Variables are important:  


## See which Variables are important:  

Model Total Profits

```{r}
profitsaux <- aggr(data, 'Month')[, 2]
profits = aggr(data, 'Month')[, 2]

h = 6
train = subset(profits, end = length(profits) - h)
test = subset(profits, start = length(profits) - h + 1)

fit = auto.arima(train, 
                 approximation = FALSE, 
                 stepwise = FALSE)
fcst = forecast(fit, h = h)

fit
autoplot(train) + autolayer(fcst) + autolayer(test) + autolayer(fit$fitted)
(rmse_profit = accuracy(fcst, test)[2, 2])
```

Take advantage of inherent structure of data: see if information within the variables is relevant for predicting total sales, or only add noise: 

```{r}
pred_profits = function(data, filtr){
  Profit_Region = aggr(data, 'Month', filtr)
  Profit_Region = Profit_Region[, (ncol(Profit_Region)/2 + 1):ncol(Profit_Region)]
  colnames(Profit_Region) = substr(colnames(Profit_Region), 3, nchar(as.character(colnames(Profit_Region))))
  
  h = 6
  train = ts(Profit_Region[1:(nrow(Profit_Region) - h), ], start = c(2015, 1), frequency = 12)
  test = ts(Profit_Region[(nrow(Profit_Region) - h + 1):nrow(Profit_Region), ], start = c(2018, 7), frequency = 12)
  
  ht_train = hts(train)
  ht_test = hts(test)
  
  f1 = forecast(ht_train, method = 'bu', fmethod = 'arima', h = h, keep.fitted = TRUE, keep.resid = TRUE)
  fcst = aggts(f1, levels = 0:1)
  
  return(accuracy(fcst[, 1], aggts(ht_test, levels = 0))[2])
}
```

```{r}
filtr = c('Segment', 'State', 'Region', 'Category', 'Sub.Category')
rmse_profits = rep(NA, length(filtr))
for (i in 1:length(filtr)){
  rmse_profits[i] = pred_profits(data, filtr[i])
}

rmse_profits
best_rmse <- rmse_profits[1]
#Best RMSE = 2222.353
```

Dynamic Harmonic Regression 
```{r}
bestfit <- list(aicc=Inf)
for(K in seq(6)) {
  fit <- auto.arima(train, xreg=fourier(train, K=K),seasonal=FALSE)
  if(fit[["aicc"]] < bestfit[["aicc"]]) {
    bestfit <- fit
    bestK <- K
  }
}

message('Number of K fourier pairs:\t', bestK)
dhr <- auto.arima(train, xreg=fourier(train, K=bestK),seasonal=FALSE)

fc_dhr <- forecast(bestfit, xreg = fourier(train, K = bestK, h = h))

accuracy(fc_dhr, test)[2,2]
accuracy(fc_dhr, test)[2,2]<best_rmse

#RMSE = 2230.580
```

TBATS model
```{r}
fit_tbats <- tbats(train)
fc_tbats <- forecast(fit_tbats, h=h)
accuracy(fc_tbats, test)[2,2]
accuracy(fc_tbats, test)[2,2]<best_rmse

#RMSE = 2775.121
```

ETS
```{r}
fit_ets <- ets(train)
fc_ets <- forecast(fit_ets, h=h)
accuracy(fc_ets, test)[2,2]
accuracy(fc_ets, test)[2,2]<best_rmse
```

STL+ETS
```{r}
windows <- expand.grid(twindow = seq(7, 35, 2), swindow = seq(7, 45, 2), RMSE = NA)

#Go through each pairs of values and store p-value and RMSE
for(row.i in 1:nrow(windows)) {
  
  #Set t and s windows for current iteration 
  twindow.i <- windows$twindow[row.i]
  swindow.i <- windows$swindow[row.i]
  
  #Get STL decomposition, setting boxcox transformation automatically
  currentSTL = stlf(train, t.window = twindow.i , s.window = swindow.i)
  
  RMSE.i<-accuracy(currentSTL, test)[2,2]
  
  #Populate p-value and RMSE in windows   
  windows$RMSE[row.i] <- RMSE.i
}


#Get white noise windows 

minerror<-which.min(windows[,"RMSE"])
best<-windows[minerror,]
best$RMSE
best$RMSE<best_rmse
```

STL+ARIMA
```{r}
windows <- expand.grid(twindow = seq(7, 35, 2), swindow = seq(7, 45, 2), RMSE = NA)

#Go through each pairs of values and store p-value and RMSE
for(row.i in 1:nrow(windows)) {
  
  #Set t and s windows for current iteration 
  twindow.i <- windows$twindow[row.i]
  swindow.i <- windows$swindow[row.i]
  
  #Get STL decomposition, setting boxcox transformation automatically
  currentSTL = stlf(train, t.window = twindow.i , s.window = swindow.i, method = 'arima')
  
  RMSE.i<-accuracy(currentSTL, test)[2,2]
  
  #Populate p-value and RMSE in windows   
  windows$RMSE[row.i] <- RMSE.i
}


#Get white noise windows 

minerror<-which.min(windows[,"RMSE"])
best<-windows[minerror,]
best$RMSE
best$RMSE<best_rmse
```

ARFIMA
```{r}
fit_arfima <- arfima(train, 
                     lambda = 'auto',
                     stepwise = FALSE, 
                     approximation = FALSE)
fc_arfima <- forecast(fit_arfima, h=h)

accuracy(fc_arfima, test)[2,2]
```

- Predicting Profit improves when considering: 
  - Segment
  - State
  - Sub Categories
  
This means that there is valuable information within these categories. When aggregated into total profits this information might be lost. Forecasts appear to decrease because they might belong to signifncantly different stochastic processes. 

We cannot control the Segment variable as it is dependent. Try to gain insight into State and Cub Categories


The best forecast is given by using a bottom-up approach with Segment. Dynamic harmonic regression does really well too, has almost the same RMSE.


Plot forecasts of profit before rightsizing 
```{r}
fc_pre <- forecast(bestfit, xreg = fourier(train, K = bestK, h = h))

autoplot(fc_dhr) + 
  xlab('Years') +
  ylab('Net Profit ($)') +
  ggtitle('')
```



Model Total Profits after filtering

```{r}
filtered = read.csv("./data_trans.csv")
filtered = preprocess2(filtered)
```


```{r}
profits = aggr(filtered, 'Month')[, 2]
autoplot(profits)

h = 6
train = subset(profits, end = length(profits) - h)
test = subset(profits, start = length(profits) - h + 1)

fit = auto.arima(train, 
                 approximation = FALSE, 
                 stepwise = FALSE)
fcst = forecast(fit, h = h)


autoplot(train) + autolayer(fcst) + autolayer(test) + autolayer(fit$fitted)
(rmse_profit = accuracy(fcst, test)[2, 2])
```

Take advantage of inherent structure of data: see if information within the variables is relevant for predicting total sales, or only add noise: 


```{r}
filtr = c('Segment', 'State', 'Region', 'Category', 'Sub.Category')
rmse_profits = rep(NA, length(filtr))
for (i in 1:length(filtr)){
  rmse_profits[i] = pred_profits(filtered, filtr[i])
}

rmse_profits
```

Dynamic Harmonic Regression 
```{r}
bestfit <- list(aicc=Inf)
for(K in seq(6)) {
  fit <- auto.arima(train, xreg=fourier(train, K=K),seasonal=FALSE)
  if(fit[["aicc"]] < bestfit[["aicc"]]) {
    bestfit <- fit
    bestK <- K
  }
}

message('Number of K fourier pairs:\t', bestK)
dhr <- auto.arima(train, xreg=fourier(train, K=bestK),seasonal=FALSE)

fc_dhr2 <- forecast(bestfit, xreg = fourier(train, K = bestK, h = h))
autoplot(fc_dhr2)
accuracy(fc_dhr2, test)[2,2]
accuracy(fc_dhr2, test)[2,2]<best_rmse
```

TBATS model
```{r}
fit_tbats2 <- tbats(train)
fc_tbats2 <- forecast(fit_tbats2, h=h)
accuracy(fc_tbats2, test)[2,2]
accuracy(fc_tbats2, test)[2,2]<best_rmse
```

ETS
```{r}
fit_ets2 <- ets(train)
fc_ets2 <- forecast(fit_ets2, h=h)
accuracy(fc_ets2, test)[2,2]
accuracy(fc_ets2, test)[2,2]<best_rmse
```

STL+ETS
```{r}
windows <- expand.grid(twindow = seq(7, 35, 2), swindow = seq(7, 45, 2), RMSE = NA)

#Go through each pairs of values and store p-value and RMSE
for(row.i in 1:nrow(windows)) {
  
  #Set t and s windows for current iteration 
  twindow.i <- windows$twindow[row.i]
  swindow.i <- windows$swindow[row.i]
  
  #Get STL decomposition, setting boxcox transformation automatically
  currentSTL = stlf(train, t.window = twindow.i , s.window = swindow.i, h = h)
  
  RMSE.i<-accuracy(currentSTL, test)[2,2]
  
  #Populate p-value and RMSE in windows   
  windows$RMSE[row.i] <- RMSE.i
}


#Get white noise windows 

minerror<-which.min(windows[,"RMSE"])
best<-windows[minerror,]
best
best$RMSE<best_rmse
```

STL+ARIMA
```{r}
windows <- expand.grid(twindow = seq(7, 35, 2), swindow = seq(7, 45, 2), RMSE = NA)

#Go through each pairs of values and store p-value and RMSE
for(row.i in 1:nrow(windows)) {
  
  #Set t and s windows for current iteration 
  twindow.i <- windows$twindow[row.i]
  swindow.i <- windows$swindow[row.i]
  
  #Get STL decomposition, setting boxcox transformation automatically
  currentSTL = stlf(train, t.window = twindow.i , s.window = swindow.i, method = 'arima', h = h)
  
  RMSE.i<-accuracy(currentSTL, test)[2,2]
  
  #Populate p-value and RMSE in windows   
  windows$RMSE[row.i] <- RMSE.i
}


#Get white noise windows 

minerror<-which.min(windows[,"RMSE"])
best<-windows[minerror,]
best$RMSE
best$RMSE<best_rmse
```

ARFIMA
```{r}
fit_arfima2 <- arfima(train, 
                     lambda = 'auto',
                     stepwise = FALSE, 
                     approximation = FALSE)
fc_arfima2 <- forecast(fit_arfima2, h=h)

accuracy(fc_arfima2, test)[2,2]
```

Plot best forecast after rightsizing
```{r}
fc <- stlf(train, t.window = 31 , s.window = 9, h = h)
clrs <- c("blueviolet", "red")

autoplot(train) +
  autolayer(fc, series="After Rightsizing", PI = TRUE) +
  autolayer(fc_pre, series='Before Rightsizing', PI = FALSE) +
  ggtitle("") +
  xlab("Year") +
  ylab("Net Profit ($)") +
  guides(colour=guide_legend(title="Data series"), 
         fill=guide_legend(title="Prediction interval")) +
  scale_color_manual(values=clrs)
```


## State

```{r}
states = aggr(data, 'Month', 'State')
states = states[, (ncol(states)/2 + 1):ncol(states)]
colnames(states) = substr(colnames(states), 3, nchar(as.character(colnames(states))))

h = 6
train = ts(states[1:(nrow(states) - h), ], start = c(2016, 1), frequency = 12)
test = ts(states[(nrow(states) - h + 1):nrow(states), ], start = c(2019, 7), frequency = 12)

ht_train = hts(train)
ht_test = hts(test)

f1 = forecast(ht_train, method = 'bu', fmethod = 'arima', h = h, keep.fitted = TRUE, keep.resid = TRUE)
fcst = aggts(f1, levels = 0:1)

autoplot(aggts(ht_train, levels = 0)) + autolayer(aggts(ht_test, levels = 0)) + autolayer(fcst[, 1])
accuracy(fcst[, 1], aggts(ht_test, levels = 0))[2]
```

```{r}
mask = apply(states, 2, sum) < 0
drop_state = colnames(states)[mask]
drop_state
```

10 unprofitabble states. 

## Sub.Categories

```{r}
categories = aggr(data, 'Month', 'Sub.Category')
categories = categories[, (ncol(categories)/2 + 1):ncol(categories)]
colnames(categories) = substr(colnames(categories), 3, nchar(as.character(colnames(categories))))

h = 6
train = ts(categories[1:(nrow(categories) - h), ], start = c(2016, 1), frequency = 12)
test = ts(categories[(nrow(categories) - h + 1):nrow(categories), ], start = c(2019, 7), frequency = 12)

ht_train = hts(train)
ht_test = hts(test)

f1 = forecast(ht_train, method = 'bu', fmethod = 'arima', h = h, keep.fitted = TRUE, keep.resid = TRUE)
fcst = aggts(f1, levels = 0:1)

autoplot(aggts(ht_train, levels = 0)) + autolayer(aggts(ht_test, levels = 0)) + autolayer(fcst[, 1])
accuracy(fcst[, 1], aggts(ht_test, levels = 0))[2]
```

```{r}
mask = apply(categories, 2, sum) < 0
drop_scat = colnames(categories)[mask]
```

3 unprofitable subcategories. 

```{r}
mask_state = ! (data$State  %in% drop_state) 
mask_cat = ! (data$Sub.Category  %in% drop_scat)
sum(mask_state | mask_cat)
sum(mask_state & mask_cat)
```

We cannot filter out these parts of the market because they represent a large propotion of the total market size of our retail store. Also they depend on each other: 

Florida: 

```{r}
fff = aggr(data, 'Quarter', c('Category', 'State'))
mask = str_split(colnames(fff), '_', simplify = TRUE)[, 3] %in% drop_state

fff = fff[, mask]
florida = fff[, c(1:3, (ncol(fff)/2 + 1):(ncol(fff)/2 + 3))]
colnames(florida) = str_split(colnames(florida), '_', simplify = TRUE)[, 2]

autoplot(window(florida[, 4:6], start = 2019)) + 
  labs(title = 'Florida', y = 'Profits', color = 'Category') +
  theme(legend.position = 'bottom')
```

Therefore, we need to analyze them together:

## Filter by Both: 

```{r}
both = aggr(data, 'Quarter', c('State', 'Sub.Category'))
both = both[, (ncol(both)/2 + 1):ncol(both)]
colnames(both) = substr(colnames(both), 3, nchar(as.character(colnames(both))))

mask = apply(both, 2, sum) < 0
drop_both = colnames(both)[mask]
split = str_split(drop_both, '_')

sum(mask)/ncol(both)
```

9% of the markets we are investing in (State and Product type) are non profitable. Loose these: 

```{r}
mask = vector(length = nrow(data))
for (i in 1:nrow(data)){
  remove = F
  for (j in 1:length(split)){
    remove = remove | ((data$State[i] == split[[j]][1]) & (data$Sub.Category[i] == split[[j]][2]))
  }
  mask[i] = remove
}

data_trans = data[! mask, ]
```

Map the outcome: 

```{r}
split2 = str_split(drop_both, '_', simplify = TRUE)
t = table(split2[,1], split2[, 2])
counts = apply(t, 1, sum)

res = data.frame(state = unique(as.character(data$State)))
other = data.frame(state = names(counts), 
                   Count = unname(counts))
res = merge(res, other, by = 'state', all.x = TRUE)
res[is.na(res)] = 0

res$Count = ifelse(res$Count == 1, 2, res$Count)
res$Count = ifelse(res$Count == 9, 8, res$Count)

plot_usmap(data = res, values = "Count", color = "white") + 
  scale_fill_continuous( low = "slategray1", high = "dodgerblue1", name = "Dropped (2021)", label = scales::comma) + 
  theme(legend.position = "right") 
```
